---
title: "R Notebook"
output: html_notebook
---
In order to color the plots (and also choose the correct size for each vertex), we will first define the following functions:

```{r}
range01 <- function(x)(x-min(x))/diff(range(x))
cRamp <- function(x){
  cols <- colorRamp(rainbow(7))(range01(x))
  apply(cols, 1, function(xt)rgb(xt[1], xt[2], xt[3], maxColorValue=255))
} 

listintolengths <- function(x){
  a <- c()
  for(item in x){
    a <- c(a, length(item))
  }
  return(a)
}

#meanforcolor <- function(map, data, coloringAPoint = -1){
#  allpoints <- data$PTS
#  auxpoints <- map$points_in_vertex
#  answer <- c()
#  for(item in auxpoints){
#    newitem <- unlist(item)
#    if(coloringAPoint == -1 || (sum(item ==coloringAPoint) == 0)){
#      num <- mean(allpoints[newitem])
#    }
#    else{

#    }
      
#    answer <- c(answer, num)
#  }
#  return(answer)
#}
meanforcolor <- function(map, data){
  allpoints <- data$PTS
  auxpoints <- map$points_in_vertex
  answer <- c()
  for(item in auxpoints){
    newitem <- unlist(item)
    num <- mean(allpoints[newitem])
    answer <- c(answer, num)
  }
  return(answer)
}


## This function computes and plots the mapper graph
MapperAndPlot <- function(distance_matrix, filter_values,
                          num_intervals, percent_overlap, num_bins) {
  map    <- mapper2D(
              distance_matrix = distance_matrix, 
              filter_values = filter_values,
              num_intervals = num_intervals,#20,20 or 30,30
              percent_overlap = percent_overlap,
              num_bins_when_clustering = num_bins # Explaining this parameter:
                # The cluster phase here is a hierarquical cluster. to define the
                # number of cluster from the hTree, we splite the real interval 
                # [0,max(height)] into num_bins_... intervals of the same size (that is 
                # if num_b... is 2, 2 intervals will be created, [0,max(height)/2) and 
                # [max(height)/2,max(height)]). We them take the clusters from the first 
                # interval, in deacrising order, that has no points (players) in it.
        )
  
  #plot
  g <- graph.adjacency(map$adjacency, mode="undirected")
  
  V(g)$color <- cRamp(meanforcolor(map,df))
  V(g)$size <- listintolengths(map$points_in_vertex)
  
  plot(g, layout = layout.auto(g), vertex.label = NA)
  #plot(g, layout = layout.auto(g), vertex.label.cex = .4)
}

```

I started with data non normalized by minutes played (the one who gave better results in PCA)
```{r}
library(igraph)
library(TDAmapper)
library(fastcluster)
library(FactoMineR)
library(ggplot2)

getwd()

#import data
df <- read.csv("../../Data/NBA-data-With-Positions.csv")

# labels
label <- df$PosicAbbrev
#levels(label) <- c(5,4,1,3,2)

# set index
rownames(df) <- df[,'Player']

# features selection -> new smaller dataset -> distance matrix
chosen_features = c('REB','AST','TOV','STL','BLK','PF','PTS')
X <- as.matrix(df[chosen_features])

normX <- t(t(X)/sqrt(diag(var(X)))) # dividing columns by the variance
                                    # (normalizing without centering)
d <- dist(normX,method = "euclidian") # This way I get what Alagapin called
                                      # "Variance normalized euclidean distance"

#Taking filter function from PCA
myPCA <- PCA(X,graph = F)
plot.PCA(myPCA,axes = c(1,2),choix = "var")
plot.PCA(myPCA,axes = c(1,2),choix = "ind", col.ind = label, label = "ind.sup")

filter <- list(matrix(myPCA$ind$coord[,1]),matrix(myPCA$ind$coord[,2]))

#cluster_cutoff_at_first_empty_bin(1,1,1)

map    <- mapper2D(
          distance_matrix = d, 
          filter_values = filter,
          num_intervals = c(30,30),#20,20 or 30,30
          percent_overlap = 30,
          num_bins_when_clustering = 4 # TODO understand and tune. Has to do with 
                                       # the cut the hierarchical clustering tree
      )

#plot
g <- graph.adjacency(map$adjacency, mode="undirected")
#TODO
# learn how to color the graph and decide a way to do it
V(g)$color <- cRamp(meanforcolor(map,df))
V(g)$size <- listintolengths(map$points_in_vertex)

plot(g, layout = layout.auto(g), vertex.label = NA)
#plot(g, layout = layout.auto(g), vertex.label.cex = .4)
```

```{r}
overlaps <- c(5,15,30,40,50,60)  
num_Intervals_array <- c(10,20,25,30,35,40)
num_bins_array <- c(2,4,6,8,10,12)

for(overlap in overlaps){
  # Open file
  filename <- paste('VarNormalized-NotDivByMin-overlapOf ',overlap,'.pdf', sep = "")
  pdf(filename)
  
  # Create frame
  par(mfrow = c(length(num_Intervals_array),
                length(num_bins_array)),
      mar = c(1, 1, 1, 1)+0.1) # 10 numbers of intervals, 15 choices of num_bins...
  
  #plot in the frame
  for (nint in num_Intervals_array){
    for(nbins in num_bins_array){
      MapperAndPlot(distance_matrix = d,
                    filter_values = filter,
                    num_intervals = c(nint,nint),
                    percent_overlap = overlap,
                    num_bins = nbins
      )
    }
  }
  #close file
  dev.off()
}
```

Same thing for normilized data (divided per minutes played)
```{r}
#import data
df <- read.csv("../../Data/NBA-data-With-Positions.csv")

# labels
label <- df$PosicAbbrev
#levels(label) <- c(5,4,1,3,2)

# features selection -> new smaller dataset -> distance matrix
chosen_features = c('REB','AST','TOV','STL','BLK','PF','PTS')
X <- as.matrix(df[chosen_features])
X <- X/df$MIN ## normalizing according to total time played

normX <- t(t(X)/sqrt(diag(var(X)))) # dividing columns by the variance
                                    # (normalizing without centering)
d <- dist(normX,method = "euclidian") # This way I get what Alagapin called
                                      # "Variance normalized euclidean distance"

#Taking filter function from PCA
myPCA <- PCA(X,graph = F)
plot.PCA(myPCA,axes = c(1,2),choix = "var")
plot.PCA(myPCA,axes = c(1,2),choix = "ind", col.ind = label, label = "ind.sup")

filter <- list(matrix(myPCA$ind$coord[,1]),matrix(myPCA$ind$coord[,2]))

#cluster_cutoff_at_first_empty_bin(1,1,1)
```

Apply mapper and plot graph
```{r}
map    <- mapper2D(
          distance_matrix = d, 
          filter_values = filter,
          num_intervals = c(30,30),#20,20 or 30,30
          percent_overlap = 50,
          num_bins_when_clustering = 5 # Explaining this parameter:
              # The cluster phase here is a hierarquical cluster. to define the
              # number of cluster from the hTree, we splite the real interval 
              # [0,max(height)] into num_bins_... intervals of the same size (that is 
              # if num_b... is 2, 2 intervals will be created, [0, max(height)/2) and 
              # [max(height)/2,max(height)]). We them take the clusters from the first 
              # interval, in deacrising order, that has no points (players) in it.
      )

#plot
g <- graph.adjacency(map$adjacency, mode="undirected")

V(g)$color <- cRamp(meanforcolor(map,df))
V(g)$size <- listintolengths(map$points_in_vertex)

plot(g, layout = layout.auto(g), vertex.label = NA)
#plot(g, layout = layout.auto(g), vertex.label.cex = .4)
```

Applying Mapper to many different configurations, and outputting them to file
```{r}
overlaps <- c(5,15,30,40,50,60)  
num_Intervals_array <- c(10,20,25,30,35,40)
num_bins_array <- c(2,4,6,8,10,12)

for(overlap in overlaps){
  # Open file
  filename <- paste('VarNormalized-DividedByMin-overlapOf ',overlap,'.pdf', sep = "")
  pdf(filename)
  
  # Create frame
  par(mfrow = c(length(num_Intervals_array),
                length(num_bins_array)),
      mar = c(1, 1, 1, 1)+0.1) # 10 numbers of intervals, 15 choices of num_bins...
  
  #plot in the frame
  for (nint in num_Intervals_array){
    for(nbins in num_bins_array){
      MapperAndPlot(distance_matrix = d,
                    filter_values = filter,
                    num_intervals = c(nint,nint),
                    percent_overlap = overlap,
                    num_bins = nbins
      )
    }
  }
  #close file
  dev.off()
}

```






Same thing for normilized data (divided per minutes played) and metrics = normalized euclidian
```{r}
#import data
df <- read.csv("../../Data/NBA-data-With-Positions.csv")

# labels
label <- df$PosicAbbrev
#levels(label) <- c(5,4,1,3,2)

# features selection -> new smaller dataset -> distance matrix
chosen_features = c('REB','AST','TOV','STL','BLK','PF','PTS')
X <- as.matrix(df[chosen_features])
X <- X/df$MIN ## normalizing according to total time played

normX <- X/apply(X,1,function(x) sqrt(sum(x^2)))
d <- dist(normX,method = "euclidian")
#normX <- t(t(X)/sqrt(diag(var(X)))) # dividing columns by the variance
#                                    # (normalizing without centering)
#d <- dist(normX,method = "euclidian") # This way I get what Alagapin called
#                                      # "Variance normalized euclidean distance"

#Taking filter function from PCA
myPCA <- PCA(normX,graph = F)
plot.PCA(myPCA,axes = c(1,2),choix = "var")
plot.PCA(myPCA,axes = c(1,2),choix = "ind", col.ind = label, label = "ind.sup")

filter <- list(matrix(myPCA$ind$coord[,1]),matrix(myPCA$ind$coord[,2]))

#cluster_cutoff_at_first_empty_bin(1,1,1)
```

Apply mapper and plot graph
```{r}
map    <- mapper2D(
          distance_matrix = d, 
          filter_values = filter,
          num_intervals = c(30,30),#20,20 or 30,30
          percent_overlap = 50,
          num_bins_when_clustering = 5 # Explaining this parameter:
              # The cluster phase here is a hierarquical cluster. to define the
              # number of cluster from the hTree, we splite the real interval 
              # [0,max(height)] into num_bins_... intervals of the same size (that is 
              # if num_b... is 2, 2 intervals will be created, [0, max(height)/2) and 
              # [max(height)/2,max(height)]). We them take the clusters from the first 
              # interval, in deacrising order, that has no points (players) in it.
      )

#plot
g <- graph.adjacency(map$adjacency, mode="undirected")

V(g)$color <- cRamp(meanforcolor(map,df))
V(g)$size <- listintolengths(map$points_in_vertex)

plot(g, layout = layout.auto(g), vertex.label = NA)
#plot(g, layout = layout.auto(g), vertex.label.cex = .4)
```

Applying Mapper to many different configurations, and outputting them to file
```{r}

overlaps <- c(5,15,30,40,50,60)  
num_Intervals_array <- c(10,20,25,30,35,40)
num_bins_array <- c(2,4,6,8,10,12)
for(overlap in overlaps){
  # Open file
  filename <- paste('UnitNormalized-DividedByMin-overlapOf ',overlap,'.pdf', sep = "")
  pdf(filename)
  
  # Create frame
  par(mfrow = c(length(num_Intervals_array),
                length(num_bins_array)),
      mar = c(.1, .1, .1, .1)) # 10 numbers of intervals, 15 choices of num_bins...
  
  #plot in the frame
  for (nint in num_Intervals_array){
    for(nbins in num_bins_array){
      MapperAndPlot(distance_matrix = d,
                    filter_values = filter,
                    num_intervals = c(nint,nint),
                    percent_overlap = overlap,
                    num_bins = nbins
      )
    }
  }
  #close file
  dev.off()
}
3
```








Same thing for all non normilized data and metrics = Variance normalized euclidian
```{r}
#import data
df <- read.csv("../../Data/NBA-data-With-Positions.csv")

# labels
label <- df$PosicAbbrev
#levels(label) <- c(5,4,1,3,2)

# features selection -> new smaller dataset -> distance matrix
chosen_features = c(4:7,9:29,31)
X <- as.matrix(df[chosen_features])
#X <- X/df$MIN ## normalizing according to total time played

#normX <- X/apply(X,1,function(x) sqrt(sum(x^2)))
#d <- dist(normX,method = "euclidian")
normX <- t(t(X)/sqrt(diag(var(X)))) # dividing columns by the variance
                                    # (normalizing without centering)
d <- dist(normX,method = "euclidian") # This way I get what Alagapin called
                                      # "Variance normalized euclidean distance"

#Taking filter function from PCA
myPCA <- PCA(normX,graph = F)
plot.PCA(myPCA,axes = c(1,2),choix = "var")
plot.PCA(myPCA,axes = c(1,2),choix = "ind", col.ind = label, label = "ind.sup")

filter <- list(matrix(myPCA$ind$coord[,1]),matrix(myPCA$ind$coord[,2]))

#cluster_cutoff_at_first_empty_bin(1,1,1)
```

Apply mapper and plot graph
```{r}
map    <- mapper2D(
          distance_matrix = d, 
          filter_values = filter,
          num_intervals = c(45,45),#20,20 or 30,30
          percent_overlap = 50,
          num_bins_when_clustering = 8 # Explaining this parameter:
              # The cluster phase here is a hierarquical cluster. to define the
              # number of cluster from the hTree, we splite the real interval 
              # [0,max(height)] into num_bins_... intervals of the same size (that is 
              # if num_b... is 2, 2 intervals will be created, [0, max(height)/2) and 
              # [max(height)/2,max(height)]). We them take the clusters from the first 
              # interval, in deacrising order, that has no points (players) in it.
      )

#plot
g <- graph.adjacency(map$adjacency, mode="undirected")

V(g)$color <- cRamp(meanforcolor(map,df))
V(g)$size  <- listintolengths(map$points_in_vertex)
(V(g)$size )

plot(g, layout = layout.auto(g), vertex.label = NA)
```

Applying Mapper to many different configurations, and outputting them to file
```{r}

overlaps <- c(5,15,30,40,50,60)  
num_Intervals_array <- c(10,20,30,35,40,45,50,55,60)
num_bins_array <- c(2,4,6,8,10,12)
for(overlap in overlaps){
  # Open file
  filename <- paste('AllFeat-VarNormalized-nonDivByMin-overlapOf',overlap,'.pdf', sep = "")
  pdf(filename)
  
  # Create frame
  par(mfrow = c(length(num_Intervals_array),
                length(num_bins_array)),
      mar = c(.1, .1, .1, .1)) # 10 numbers of intervals, 15 choices of num_bins...
  
  #plot in the frame
  for (nint in num_Intervals_array){
    for(nbins in num_bins_array){
      MapperAndPlot(distance_matrix = d,
                    filter_values = filter,
                    num_intervals = c(nint,nint),
                    percent_overlap = overlap,
                    num_bins = nbins
      )
    }
  }
  #close file
  dev.off()
}
3
```



This function recieves the cluster label and returns the players that are in that cluster
```{r}
clusterintoplayers <- function(data, map, label){
  players <- map$points_in_vertex[label]
  answer <- c()
  allplayers <- data$Player
  for(player in players){
    answer <- c(answer, toString(allplayers[player]))
  }
  return (answer)
}

#example:
clusterintoplayers(df, map, 100)
```


